Designing Large-Scale Imaging Surveys for a Retrospective Relative Photometric Calibration

### 2012-03-19 paper/self_calibration_I.tex
euclid: minor english editing to paper
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@20439 03a3cea6-2c03-0410-ac9b-9271b1c66c29

\begin{abstract}
We study the self-calibration---determination of complex system
response from science data alone---for precise photometric catalogs
from wide-field imaging surveys. We create an artificial sky of stars
and synthetically observe it under four basic survey strategies,
creating an end-to-end simulation of an imaging survey for each. These
catalog-level simulations include realistic measurement uncertainties
and a complex, position-dependent instrument response. We then use a
self-calibration technique to recover the instrument response by
simultaneously fitting for all the star fluxes and the parameters of a
position-dependent flat-field, optimizing a justified likelihood
function.  For realism, we deliberately fit with a wrong noise model
and a flat-field functional basis that does not include the model that
generated the synthetic data.  We demonstrate that with a favorable
survey strategy, a complex (but smooth) instrument response can be
precisely self-calibrated. We also find that returning the same
sources to very different focal plane positions is the key property of
any survey strategy designed for accurate retrospective
calibration. The results of this work suggest the following advice for
those considering the design of large-scale imaging surveys: Do not
use a regular tiling of the sky; instead return the same sources to
very different focal plane positions.
\end{abstract}

@@

In this paper, we argue that the next generation of large-scale imaging surveys should have their observation strategies designed from the very start with this kind of ``self-calibration'' in mind. This work focuses on the relative photometric calibration of a typical imaging instrument only, although similar techniques could be used to constrain many other calibration parameters, such as the optical distortion, point-spread-functions and so-on. The self-calibration technique utilizes the multiple measurements of (non-varying) sources, at different focal plane positions and at different times within the survey, to constrain the relative instrument response by requiring that the post-calibration measurements yield consistent flux values. Through end-to-end, catalog-level survey simulations, we aim to identify the important properties of survey strategies that makes them advantageous for this kind of self-calibration.

@@

In this work we do not produce pixelated images.  Instead we create catalog-level simulations with realistic measurement uncertainties, including unmodeled stellar variability.  Complex effects are included in the simulations that are (deliberately) not correctly modeled at the analysis stage, in order to simulate the effects of unknown systematic errors within the catalog.

@@

\subsection{A Single Exposure}
\label{sec:single_exposure}
With a camera pointing $(\alpha, \beta)$ and camera orientation $\theta$, we find the focal plane positions of the sources on the synthetic sky that fall within the instrument's field-of-view. In our simulations, we use a large instrument field-of-view of $0.76~ \text{deg} \times 0.72~\text{deg}$; a size consistent with the next generation of large survey imagers. An example of a single pointing exposure is shown in Figure \ref{fig:single_image}.

\subsubsection{Measured Count Rates}
To generate photometric catalogs, the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position-dependent instrument response model $f_\true$ and a measurement noise model. For a measurement $i$ the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$, which is a function of focal plane position $\vec{x_i}$, and the source's true flux $s_{k,\true}$ ,
\begin{eqnarray*}
c_i = f_{\true}(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} + e_{i} \quad ,
\end{eqnarray*}
where $\vec{q}_\true$ are the parameters defining the \textit{true} instrument response (see Section \ref{sec:instrument_response_model}), and $e_i$ is a noise contribution drawn from the Normal Distribution $N(e|0,{\sigma_\true}^2)$.

@@

With the Survey C and D catalogs, the instrument response is accurately recovered. As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top), the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~percent$ level. That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the model basis used. The residuals in this comparison are low; this can also be clearly seen in the factor of $\sim 10$ improvement in the best-in-basis badness $B_\basis$ compared to the true badness $B_\true$. The results from these two survey catalogs are clear: if the correct redundancy is built into the survey, then self-calibration calibration can be used to accurately constrain the relative instrument response, with presumably the majority of the remaining errors coming from the limitations in the model used to represent the instrument response.

>>>>

\begin{abstract}
We study the self-calibration -- the determination of a complex system
response from science data alone -- for precise photometric catalogs
from wide-field imaging surveys. We create an artificial sky of sources
and synthetically observe it under four basic survey strategies,
creating an end-to-end simulation of an imaging survey for each. These
catalog-level simulations include realistic measurement uncertainties
and a complex, position-dependent instrument response. We then use a
self-calibration technique to recover the instrument response by
simultaneously fitting for all the star fluxes and the parameters of a
position-dependent flat-field, optimizing a justified likelihood
function.  For realism, we deliberately fit with a wrong noise model
and a flat-field functional basis that does not include the model that
generated the synthetic data.  We demonstrate that with a favorable
survey strategy, a complex (but smooth) instrument response can be
precisely self-calibrated. We also find that returning the same
sources to very different focal plane positions is the key property of
any survey strategy designed for accurate retrospective
calibration. The results of this work suggest the following advice for
those considering the design of large-scale imaging surveys: Do not
use a regular tiling of the sky; instead return the same sources to
very different focal plane positions.
\end{abstract}

@@

In this paper, we argue that the next generation of large-scale imaging surveys should have their observation strategies designed from the very start with this kind of ``self-calibration'' in mind. This work focuses on the relative photometric calibration of a typical imaging instrument only, although similar techniques could be used to constrain many other calibration parameters, such as the optical distortion, point-spread-functions and so-on. The self-calibration technique utilizes the multiple measurements of sources, at different focal plane positions and at different times within the survey, to constrain the relative instrument response by requiring that the post-calibration measurements yield consistent flux values. Through end-to-end, catalog-level survey simulations, we aim to identify the important properties of survey strategies that makes them advantageous for this kind of self-calibration.

@@

In this work we do not produce pixelated images.  Instead we create catalog-level simulations with realistic measurement uncertainties, including unmodeled source variability.  Complex effects are included in the simulations that are (deliberately) not correctly modeled at the analysis stage, in order to simulate the effects of unknown systematic errors within the catalog.

@@

\subsection{A Single Exposure}
\label{sec:single_exposure}
With a telescope pointing $(\alpha, \beta)$ and camera orientation $\theta$, we find the focal plane positions of the sources on the synthetic sky that fall within the instrument's field-of-view. In our simulations, we use a large instrument field-of-view of $0.76~ \text{deg} \times 0.72~\text{deg}$; a size consistent with the next generation of large survey imagers. An example of a single pointing exposure is shown in Figure \ref{fig:single_image}.

\subsubsection{Measured Count Rates}
To generate photometric catalogs, the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position-dependent instrument response model $f_\true$ and a measurement noise model. For a measurement $i$, the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$ -- which is a function of focal plane position $\vec{x_i}$ -- and the source's true flux $s_{k,\true}$ ,
\begin{eqnarray*}
c_i = f_{\true}(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} + e_{i} \quad ,
\end{eqnarray*}
where $\vec{q}_\true$ are the parameters defining the \textit{true} instrument response (see Section \ref{sec:instrument_response_model}), and $e_i$ is a noise contribution drawn from the Normal Distribution $N(e|0,{\sigma_\true}^2)$.

@@

With the Survey C and D catalogs, the instrument response is accurately recovered. As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top), the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1$~percent level. That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the model basis used. The residuals in this comparison are low; this can also be clearly seen in the factor of $\sim 10$ improvement in the best-in-basis badness $B_\basis$ compared to the true badness $B_\true$. The results from these two survey catalogs are clear: if the correct redundancy is built into the survey, then self-calibration calibration can be used to accurately constrain the relative instrument response, with presumably the majority of the remaining errors coming from the limitations in the model used to represent the instrument response.
